---
title: "Assignment_3"
author: "Hanish Bhogadi"
date: '2022-10-19'
output:
  pdf_document: default
  word_document: default
---
 
1. Formulate and solve this transportation problem using R

Converting all details into table format

```{r}
tab <- matrix(c(22,14,30,600,100,
                16,20,24,625,120,
                80,60,70,"-","-") , ncol=5 , byrow=TRUE)

colnames(tab) <- c("Warehouse1","Warehouse2","Warehouse3","Prod cost","Prod Capacity") 
row.names(tab) <- c("Plant A","Plant B","Demand")
tab <- as.table(tab)
tab
```
$$\text{Min} \hspace{.2cm} TC = 622 X_{11} +614 X_{12} +630 X_{13} +641 X_{21}+ 645X_{22}+649X_{23}$$ 
 

/text{subject to}

#Production Capacity consntraints 
Production plant A :
$$ X_{11} +X_{12}+X_{13}+ \le 100 $$
Production Plant B :
$$ X_{21} +X_{22}+X_{23}+ \le 120 $$
#Demand Consntraints 

Demand Warehouse 1 :
$$ X_{11} + X_{21} \ge 80 $$
Demand Warehouse 2 :
$$ X_{12} + X_{22} \ge 60 $$
Demand Warehouse 3 :
$$ X_{13} + X_{23} \ge 70 $$
Non-negativity of the variables

$$ X_{ij} \ge 0 $$
Where 

$$ i = 1,2,3 $$
And 

$$ j = 1,2,3 $$
It is unbalanced since Demand not equal to supply so we have created the dummy row as warehouse_4

```{r}
library(lpSolveAPI)
library(lpSolve)
# Set up cost matrix
Trans_cost <- matrix(c(622,614,630,0,
                641,645,649,0) , ncol=4 , byrow=TRUE)
#defyning rows and coloumns 
colnames(Trans_cost) <- c("Warehouse_1","Warehouse_2","Warehouse_3","Dummy")
rownames(Trans_cost) <- c("Plant_A", "Plant_B")
Trans_cost
```
```{r}
#setting up constraint signs and right-hand sides(Production side)
row.signs <- rep("<=",2)
row.rhs <- c(100,120)

#Demand side constraints#
col.signs <- rep(">=",4)
col.rhs <- c(80,60,70,10)

#solve the model
lptrans <- lp.transport(Trans_cost,"min",row.signs,row.rhs,col.signs,col.rhs)

lptrans$solution
```
When solved the transportation problem, I got the values of the variables as
$$x_{12}= 60$$
$$x_{13}= 40$$
$$x_{21}= 80$$
$$x_{23}= 30$$
$$x_{24}= 10$$
```{r}
lptrans$objval
```
2)Formulate the dual of the transportation problem

As we know the primal was to minimize the transportation cost the dual of it 
would be to maximize the value
added(VA).

```{r}
cost_2 <- matrix(c(622,614,630,100,"u1",
641,645,649,120,"u2",
80,60,70,220,"-",
"v1","v2","v3","-","-"),ncol = 5,nrow = 4,byrow = TRUE)
colnames(cost_2) <- c("Warehouse_1", "Warehouse_2","Warehouse_3","Production Capacity","Supply(Dual)")
rownames(cost_2) <- c("Plant_A","Plant_B","Demand","Demand(Dual)")
```


p and q will be the variables for the dual.
$$  \text{Max} \hspace{.2cm} Z = 100 p_{1} + 120 p_{2} +80 q_{1} + 60 q_{2} + 70 q_{3}$$


Subject to the following constraints

$$  p_{1} + q_{1} \le 622 $$ 
$$  p_{1} + q_{2} \le 614 $$ 
$$  p_{1} + q_{3} \le 630 $$ 
$$  p_{2} + q_{1} \le 641 $$ 

$$  p_{2} + q_{2} \le 645 $$ 
$$  p_{2} + q_{3} \le 649 $$ 
Where
q1 = Warehouse_1

q2 = Warehouse_2

q3 = Warehouse_3

p1 = Plant_1

p2 = Plant_2
```{r}
#Objective function

f.obj <- c(100,120,80,60,70)

#transposed from the constraints matrix in the primal
f.con <- matrix(c(1,0,1,0,0,
                  1,0,0,1,0,
                  1,0,0,0,1,
                  0,1,1,0,0,
                  0,1,0,1,0,
                  0,1,0,0,1), nrow = 6, byrow = TRUE)

f.dir <- c("<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=")

f.rhs <- c(622,614,630,641,645,649)
lp("max",f.obj,f.con,f.dir,f.rhs)

lp("max",f.obj,f.con,f.dir,f.rhs)$solution
```
So Z=139,120 dollars and variables are: $$ p_{1} = 614 $$which represents Plant A
 $$ p_{2} = 633 $$which represents Plant B
$$ q_{1} = 8 $$ which represents Warehouse 1
 $$ q_{3} = 16 $$which represents Warehouse 3
 
3) Economic Interpretation of the dual

Observations:

The maximum combined costs of shipping and production will be 139,120 dollars.

There is a minimal Z=132790 (Primal) and a maximum Z=139120 (Dual).This problem is trying to find a maximum and a minimum.As a result, we discovered that we should not be shipping from Plant(A/B) to all three warehouses at the same time. We should be shipping from:

$$60 x_{12}$$ which is 60 Units from Plant A to Warehouse 2.


$$40 x_{13}$$ which is 40 Units from Plant A to Warehouse 3.

$$80 x_{21}$$which is 80 Units from Plant B to Warehouse 1.

$$30 x_{23}$$ which is 30 Units from Plant B to Warehouse 3.
We will Max the profit from each distribution to the respective capacity.

We have the following:
$$  p_{1} ^0 -q_{1} ^0 \le 622 $$ then we subtract $$  q_{1} ^0 $$ to the other side to get  $$ p_{1} ^0\le 622 - q_{1} ^0 $$
To compute it would be $614 <= (-8+622) which is correct. we would continue to evaluate these equations:
$$  p_{1}\le 622 - q_{1} => 614\le622 - 8 = 614=>correct$$
$$  p_{1}\le 614 - q_{2} => 614\le614 - 0 = 614=>correct$$
$$  p_{1}\le 630 - q_{3} => 614\le630 - 16 = 614=>correct$$
$$  p_{2}\le 641 - q_{1} => 633\le614 - 8 = 633=>correct$$
$$  p_{2}\le 645 - q_{2} => 633\le645 - 0 = 645=> Incorrect$$
$$  p_{2}\le 649 - q_{3} => 633\le649 - 16 = 633=>correct$$
From the Duality-and-Sensitivity, By updating each of the columns we can test for the shadow price. In our LP Transportation problem , we change 100 to 101 and 120 to 121.Here we can see it R.

```{r}
row.rhs1 <- c(101,120)
row.signs1 <- rep("<=",2)
col.rhs1 <- c(80,60,70,10)
col.signs1 <- rep(">=",4)
row.rhs2 <- c(100,121)
row.signs2 <- rep("<=",2)
col.rhs2 <- c(80,60,70,10)
col.signs2 <- rep(">=",4)

lp.transport(Trans_cost,"min",row.signs,row.rhs,col.signs,col.rhs)

lp.transport(Trans_cost,"min",row.signs1,row.rhs1,col.signs1,col.rhs1)
lp.transport(Trans_cost,"min",row.signs2,row.rhs2,col.signs2,col.rhs2)
```

By taking the minimum of this specific function, seeing the number go down by 19 means the shadow price is 19, which was calculated by adding 1 to every plant.The Plant B does not have a shadow price. We also

From the dual variable $$ q_{2}$$ where Marginal Revenue <= Marginal Cost. The equation was $$  p_{2}\le 645 - q_{2} => 633\le645 - 0 = 645=> Incorrect$$ and this was found by using $$  p_{1} ^0 - q_{1} ^0 \le 622 $$ then we subtract $$  q_{1} ^0 $$ to the other side to get  $$ p_{1} ^0\le 622 - q_{1} ^0 $$

```{r}
lp("max", f.obj,f.con, f.dir,f.rhs)$solution
```

$$q_{2}=0$$ .

The interpretation from above:
from the primal:
 $$60x_{12}$$which is 60 Units from Plant A to Warehouse 2.
 $$40x_{13}$$which is 40 Units from Plant A to Warehouse 3.
 $$80x_{21}$$which is 80 Units from Plant B to Warehouse 1.
 $$30x_{23}$$which is 60 Units from Plant B to Warehouse 3.
 
From the dual

The agenda is to get MR=MC.In five of the six cases, MR<=MC.Only Plant B to Warehouse 2 fails to satisfy this requirement.From the primal, we can see that no AED devices will be shipped there.
```{r}

```
